---
title: "Code part 2"
author: "Jolene Branch"
date: "July 5, 2021"
output:
  pdf_document: default
  html_document: default
---
# Description: This uses heatmap() and the corrplot library to create visualizations.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I probably had too many libraries open in the other R scipt, so sent dataset to this one to try and spread.
```{r}
install.packages("dplyr")
library(dplyr)
```

```{r import df1, rename it df2}
df2 <- read.csv('df1.csv',stringsAsFactors = FALSE)
head(df2)
```

```{r spread df2}
spread_COVID <- spread(df2, key = Type, value = Cases)
head(spread_COVID)
```

# try importing another library
```{r}
install.packages("tidyr")
install.packages("devtools")
library(tidyr)
library(devtools)

```

```{r}
spread_COVID <-
  pivot_wider(df2, names_from = Type, values_from = Cases)
head(spread_COVID)
```

# I don't know how much all that helped, but now I want to create a heatmap using just a few columns from merge_one_trimmed.  I resaved the .csv and renamed some of the columns.  Here I'll just import the ones I want to use, then convert my dataframe into a matrix

```{r}
install.packages("ggplot2")
install.packages("plotrix")
library(ggplot2)
library(plotrix)
```

```{r import .csv file}
data <- read.csv("for_heatmap_E_hemi_only.csv", header=TRUE)
data <- data.matrix(data[,-1])
heatmap(data, Rowv = NA,Colv = NA, col=heat.colors(256), scale = "column")
```

# That is way too many countries to be able to see anything! Try the variables compared to themselves
```{r}
heatmap(as.matrix(data),
        scale = "column",
        col = heat.colors(256),
        main = "Title Goes Here",
        Rowv = NA,
        Colv = NA)
```

# That did the same thing.  Let's try without the country IDs.
```{r}
data1 <- read.csv("E_hemi_only_no_names.csv", header = TRUE)
heatmap(as.matrix(data1),
        scale = "column",
        col = heat.colors(256),
        main = "Title Goes Here",
        Rowv = NA,
        Colv = NA)
```

# What I need is a correlation matrix, not a heatmap. 

# From http://www.sthda.com/english/wiki/correlation-matrix-an-r-function-to-do-all-you-need
```{r}
install.packages("corrplot")
library(corrplot)

```

```{r the code for this}
source("http://www.sthda.com/upload/rquery_cormat.r")
```

# Initial matrix, without converting chronic illness counts to rates.  This turned out to be very wrong!
# DO NOT RE-RUN THIS!
```{r pull into matrix}
#data3 <- read.csv("for_heatmap_E_hemi_major_countries_only.csv", header=TRUE)
#data3 <- data.matrix(data3[,-1])
```

# This is wrong, but it was my first try.  I thought I'd found a 'smoking gun!'
# DO NOT RE-RUN THIS!
```{r follow how website did it}
#rquery.cormat(data3, type = "full")
```

# After converting chronic disease COUNTS to RATES:
```{r pull into matrix, corrected death rate}
data4 <- read.csv("for_heatmap_E_hemi_major_countries_only.csv", header=TRUE)
data4 <- data.matrix(data4[,-1])
```

```{r follow how website did it, corrected death rate}
rquery.cormat(data4, type = "full")
```

```{r half a matrix}
rquery.cormat(data4, type="upper")
```


```{r heatmap}
cormat<-rquery.cormat(data4, graphType="heatmap")
```

```{r correlation matrix without the graph}
rquery.cormat(data4, graph=FALSE)
```

